---
title: "earthquake"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{earthquake}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(earthquake)
```


# Earthquake package

The functions provided by this package use data from the [U.S. National Oceanographic and Atmospheric Administration (NOAA)](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1) Significant Earthquake Database. 

## sample data

The following is the data included in the package


```{r, message=FALSE}
filename <- system.file("extdata", "earthquakes-2020-10-19_15-21-05_+0300.tsv", package = "earthquake")
earthquakes <- readr::read_delim(filename, delim = "\t")
```

## Functions for cleaning the data

There are three functions for cleaning the earthquakes data.

`eq_clean_data` adds a `DATE` column and cleans `LOCATION_NAME` by using the 
`eq_location_clean` function to create the location column and the `eq_country_clean`
to create the country column

`eq_location_clean` returns a title case character string with the location's 
country removed. 

```{r}
earthquake:::eq_location_clean("Canada:Toronto")
```

`eq_country_clean` returns a title case character string with the only the 
location's country remaining

```{r}
earthquake:::eq_country_clean("Canada:Toronto")
```

You can create a `clean_earthquakes` data frame like this:

```{r, message=FALSE}
library(dplyr)
clean_earthquakes <- earthquake:::eq_clean_data(earthquakes)
```

## Geoms for earthquake time line

This package contains two `ggplot2` geoms for plotting earthquake time lines.

### geom_timeline

`geom_timeline` plots a time line of earthquakes with a point for each earthquake. The `x` axis is a date and the `y` aesthetic is a factor (e.g. country). Optional aesthetics include `color`, `size`, and `alpha`.

Here is an example:

```{r, fig.show="hold", fig.width=7, message=FALSE}
library(ggplot2)

eq = dplyr::filter(clean_earthquakes, COUNTRY == "Mexico" & lubridate::year(DATE) >= 2000)

ggplot2::ggplot(eq,aes(x=DATE,y = COUNTRY, color = TOTAL_DEATHS, size = TOTAL_DEATHS)) +
  geom_timeline(alpha=.5)
```

### geom_timeline_label

This geom is intended to be used together with the `geom_timeline` geom to add a vertical line with an annotation for each data point on an earthquake timeline. The `x` aesthetic specifies the date, the`label` aesthetic specifies the label for the annotation.

Here is an example:

```{r, fig.width=7, fig.height=7}

ggplot2::ggplot(eq, aes(x=DATE,y = COUNTRY, color = TOTAL_DEATHS, size = TOTAL_DEATHS)) +
  geom_timeline(alpha=.5)+
  geom_timeline_label(aes(label = LOCATION_NAME, n_max = 5))
```

## Mapping functions

### eq_map

The `eq_map` function produces a map of earthquake epicenters (LATITUDE/LONGITUDE) and annotates each point with a popup window containing annotation data stored in a column of the data frame. The radius of each circle is proportional to the earthquake's magnitude (EQ_PRIMARY).

Here is an example:

```{r, fig.width=7, fig.height=7}
eq_map(eq, annot_col = "COUNTRY")
```

### eq_create_label

This function creates HTML labels with location, magnitude, and total deaths for each earthquake.

Here is an example:

```{r, fig.width=7, fig.height=7}
eq = dplyr::mutate(eq, popup_text = eq_create_label(eq))
eq_map(eq, annot_col = "popup_text")
```



